---
title: "Simple logistic regression"
author: "Amanda Wittendorf"
date: "`r Sys.Date()`"
output: pdf_document
---
# SIMPLE LOGISTIC REGRESSION 

Here the dataset *perulung.csv* is examinated. It contains
measurements from 636 Peruvian children. More specifically: 

- $\text{FEV1}$: Forceret Ekspiratorisk Volumen in the 1. secound 

- $\text{height}$: childrens height (cm) 

- $\text{age}$: childrens age (year) 

- $\text{sex}$: gender (0 = girls / 1 = boys)

- $\text{respsymptoms}$: measured respiratory symptoms within the past year (0 = no / 1 = yes)

```{r}
perulung <- read.csv("perulung.csv")
glimpse(perulung)
```

The variables $\text{sex}$ and $\text{respsymptoms}$ needs to be recoded
so it's easier to use later

```{r}
perulung <- perulung |> 
  mutate(sex = factor(sex,
                      levels = c(0, 1), 
                      labels = c("Girls", "Boys")),
         sex = fct_relevel(sex, "Boys"),
         respsymptoms = factor(respsymptoms,
                               levels = c(0, 1),
                               labels = c("No", "Yes")),
         respsymptoms = fct_relevel(respsymptoms, "Yes"))
```


## TASK A


We will continue with the dataset *perulung.csv* from the previous task 

### a. Compare the prevalence of $\text{respsymptoms}$ between the two sexes 

Because the $\text{sex}$ is a factor the output of a regression model vil only show the reference group ($\text{Intercept} = {sex}_{boys}$) and the difference between the two sexes ($\text{sex}_{girls} = {sex}_{difference}$). Therefore, an additional *linear combination* needs to bee calculated - otherwise the result for $\text{sex}_{girls}$ will not be calculated.

Compared to the previous regression model this one is going to be a logistic regression model because the outcome is a factor instead of a continouse variable. 

```{r}
# Reference group ($\text{Intercept} = {sex}_{boys}$) & difference
table_2 <- list(
  "Model 2" = glm(respsymptoms ~ sex, data = perulung, family = binomial) |> parameters(exponentiate = TRUE) |> as.data.frame()
  ) |> 
  bind_rows(.id = "Table 2: Simple logistic regression") |> 
  select("Table 2: Simple logistic regression" = Parameter, 
         Estimate = Coefficient, CI_low, CI_high, p) |> 
  mutate(
    across(c(Estimate, CI_low, CI_high), ~ round(.x, 2)),
    p = case_when(
      p < 0.001 ~ "< 0.001",
      p < 0.05  ~ "< 0.05",
      TRUE      ~ as.character(round(p, 3))
    )
  )

# Other group ($\text{sex}_{girls}$)
tabel_hyp <- list(
  "sexBoys" = hypotheses(reg_model, hypothesis = "`(Intercept)` + sexGirls = 0", df = reg_model$df.residual) |> parameters() |> as.data.frame() 
  ) |> 
  bind_rows(.id = "Table: Linear combination") |> 
  select("Table: Linear combination" = Parameter, 
         Estimate = Coefficient, SE, CI_low, CI_high, p) |> 
  mutate(
    across(c(Estimate, SE, CI_low, CI_high), ~ round(.x, 2)),
    p = case_when(
      p < 0.001 ~ "< 0.001",
      p < 0.05  ~ "< 0.05",
      TRUE      ~ as.character(round(p, 3))
    ),
    "Table: Linear combination" = case_match("Table: Linear combination",
                           "(Intercept)`+sexGirls=0" ~ "Respsymptom, boys",
    .default = "sexBoys")
    )

# Combining the two calculations
collect_list <- list(
  "Odds: Boys (Intercept)" = glm(respsymptoms ~ sex, 
                             data = perulung, family = binomial) |> parameters(exponentiate = TRUE) |> as.data.frame() |> slice(1),
  "Odds ratio: Girls vs. Boys"  = glm(respsymptoms ~ sex, 
                             data = perulung, family = binomial) |> parameters(exponentiate = TRUE) |> as.data.frame() |> slice(2),
  "Odds: Girls"  = hypotheses(reg_model, hypothesis = "`(Intercept)` + sexGirls = 0", 
                                    df = reg_model$df.residual) |> parameters(exponentiate = TRUE) |> as.data.frame()
  ) |>  bind_rows(.id = "Parameter") |> 
  select("Table 2: Simple logistic regression" = Parameter, Estimate = Coefficient, SE, CI_low, CI_high, p) |> 
  mutate(
    across(c(Estimate, SE, CI_low, CI_high), ~ round(.x, 2)),
    p = case_when(
      p < 0.001 ~ "< 0.001",
      p < 0.05  ~ "< 0.05",
      TRUE      ~ as.character(round(p, 3))
    )
  )
  
# Print 
collect_list
```
**Intercept:** Represents $\text{odds}_{boys}$ in which boys with *respiratory symptoms* are compared to boys *without respiratory symptoms*. The results show, that the odds is `r collect_list$Estimate[1]` (95% CI: [`r collect_list$CI_low[1]`; `r collect_list$CI_high[1]`], *p* `r collect_list$p[1]`). This means that among boys, the likelihood of having *respiratory symptoms* is highly statistically significant, with `r collect_list$Estimate[1]` boys experiencing symptoms for every 1 boy who does not.

**Odds: Girls:** Like the above, this represents $\text{odds}_{girls}$ in which boys with *respiratory symptoms* are compared to girls *without respiratory symptoms*. The results show, that the odds is `r collect_list$Estimate[3]` (95% CI: [`r collect_list$CI_low[3]`; `r collect_list$CI_high[3]`], *p* `r collect_list$p[3]`). This means that among girls, the likelihood of having *respiratory symptoms* is not statistically significant, with `r collect_list$Estimate[3]` girls experiencing symptoms for every 1 girl who does not.

**Odds ratio: Girls vs. Boys:** Here, the two odds, ($\text{sex}_{girls}$ divided by $\text{odds}_{boys}$, are compared (`r collect_list$Estimate[2]` (95% CI: [`r collect_list$CI_low[2]`; `r collect_list$CI_high[2]`], *p* `r collect_list$p[2]`)). It shows, that girls have `r collect_list$Estimate[2]` lower odds for *respiratory symptoms* than boys - this is not significantly. 

