---
title: "TASK (dataset: chd)"
author: "Amanda Wittendorf"
date: "`r Sys.Date()`"
output: html_document
---

## 1. DESKRIPTION ON THE DATASET

The dataset contains measurements a large American study named *The Framingham Heart Study*, whose purpose was to identify factors that could contribute to cardiovascular diseases

from 636 Peruvian children. More specifically: 

- $\text{id}$:     individuals ID-number 

- $\text{chd}$:    cardiovascular disease after 18 years of follow-up (1 = yes, 0 = no)

- $\text{sbp160}$: systolic blood pressure (1 = sbp > 160, 0 = sbp < 160)

- $\text{age}$:    individuals age (year) 

- $\text{sex}$:    gender (1 = male, 2 = female)


```{r dataset, message=FALSE, include=FALSE}
library(statistik)

chd <- read.csv("chd.csv") |> 
  mutate(sex = factor(sex,
                      levels = c(1, 2), 
                      labels = c("Male", "Female")),
         sex = fct_relevel(sex, "Male"),
         sbp = factor(sbp160,
                         levels = c(1, 0),
                         labels = c(">160", "<160")),
         sbp = fct_relevel(sbp, "<160")
         )
```


## 1. SIMPLE LOGISTIC REGRESSION MODEL

### 1.1 ASSOCIATION BETWEEN SYSTOLIC BLOOD PRESSURE ($\text{sbp160}$) & THE RISK OF GETTING CARDIOVASCULAR DISEASE AFTER 18 YEARS OF FOLLOW-UP ($\text{chd}$)

**The model**
$$chd = \beta_\text{0} + \beta_\text{1} \cdot sbp160 $$
```{r regression_resp_sex, message=FALSE, warning=FALSE, include=FALSE}
table_1 <- list(
  "Model" = glm(chd ~ sbp, data = chd, family = binomial) |> parameters(exponentiate = TRUE) |> as.data.frame()
  ) |> 
  bind_rows(.id = "Table 1: Simple logistic regression") |> 
  select("Table 1: Simple logistic regression" = Parameter, 
         Estimate = Coefficient, CI_low, CI_high, p) |> 
  mutate(
    across(c(Estimate, CI_low, CI_high), ~ round(.x, 2)),
    p = case_when(
      p < 0.001 ~ "< 0.001",
      p < 0.05  ~ "< 0.05",
      TRUE      ~ as.character(round(p, 3))
    )
  )
```
```{r table_resp_sex, echo=FALSE, message=FALSE, warning=FALSE}
library(stargazer)

## Models
model   <- glm(chd ~ sbp, data = chd, family = binomial)

## Transforming from log-odds to odds
OR_vector <- exp(coef(model))
CI_vector <- exp(confint(model))
p_values  <- summary(model)$coefficients[, 4]

## Table
stargazer(model, 
          type = "text",                  
          coef = list(OR_vector), 
          title = "Table 1: Association between cardiovascular disease and SBP (ref. = <160 mmHg)",
          ci = TRUE,                      
          ci.custom = list(CI_vector),    
          p = list(p_values),             
          covariate.labels = c("SBP (>160)", "Intercept"),
          digits = 2,
          star.cutoffs = c(0.05, 0.01, 0.001), 
          keep.stat = c("n")
)
```
**INTREPETATION**

**Intercept:** 
The statistically significant odds of developing cardiovascular disease ($\text{chd}$) for individuals with normal systolic blood pressure $\text{sbp}_{<160}$ (reference group) is `r table_1$Odds[1]` (`r table_1$CI_low[1]`; `r table_1$CI_high[1]`, p `r table_1$p[1]`). This Oddss that for every 1 individual with normal systolic blood pressure who does not develop cardiovascular disease, `r table_1$Odds[1]` individual(s) will develop the disease.

**SBP (>160):**
The statistically significant odds for a a person with a elevated systolic blood pressure ($\text{sbp}_{>160}$) has `r table_1$Odds[2]` times higher odds to get cardiovascular disease compared to those with normal blood pressure ($\text{sbp}_{<160}$).


##### 1.1.1 The corresponding odds ratio for men and women separately

To calculate the OR for the genders separately the function `filter` is included in separate regression model; one for the female, and one for the males. 

```{r regression_gender_sep, message=FALSE, warning=FALSE, include=FALSE}
table_2 <- list(
  "Model (female)" = glm(chd ~ sbp, 
                         data = filter(chd, sex == "Female"), family = binomial) |> parameters(exponentiate = TRUE) |> as.data.frame(),
  "Model (male)" = glm(chd ~ sbp, 
                 data = filter(chd, sex == "Male"), family = binomial) |> parameters(exponentiate = TRUE) |> as.data.frame()
  ) |> 
  bind_rows(.id = "Table 2: Separate regression model (genders)") |> 
  select("Table 2: Separate regression model (genders)" = Parameter, 
         Estimate = Coefficient, CI_low, CI_high, p) |> 
  mutate(
    across(c(Estimate, CI_low, CI_high), ~ round(.x, 2)),
    p = case_when(
      p < 0.001 ~ "< 0.001",
      p < 0.05  ~ "< 0.05",
      TRUE      ~ as.character(round(p, 3))
    )
  )
```
```{r comp_table2, echo=FALSE, message=FALSE, warning=FALSE}
library(stargazer)

## Models
model_female <- glm(chd ~ sbp, 
                     data = filter(chd, sex == "Female"), family = binomial)

model_male   <- glm(chd ~ sbp, 
                     data = filter(chd, sex == "Male"), family = binomial)

## ## Transforming from log-odds to odds (Både for kvinder og mænd)
OR_female <- exp(coef(model_female))
CI_female <- exp(confint(model_female))
p_female  <- summary(model_female)$coefficients[, 4]

OR_male   <- exp(coef(model_male))
CI_male   <- exp(confint(model_male))
p_male    <- summary(model_male)$coefficients[, 4]

## Table
stargazer(model_female, model_male,
          type = "text",
          title = "Table 2: Separate regression model (genders)",
          column.labels = c("Model (female)", "Model (male)"),
          coef = list(OR_female, OR_male),   
          ci = TRUE,
          ci.custom = list(CI_female, CI_male),  
          p = list(p_female, p_male),         
          covariate.labels = c("SBP (<160)", "Intercept"),
          digits = 2,
          star.cutoffs = c(0.05, 0.01, 0.001),
          keep.stat = c("n")
)
```
**INTERPRETATION**

**Intercept:** 
The same as before but separately for males and females... 

However, the odds of developing cardiovascular disease for $\text{gender}_{males}$ are not statistically significant $\text{sbp}_{<160}$ (reference group) is `r table_2$Odds[3]` (`r table_2$CI_low[3]`; `r table_2$CI_high[3]`, p `r table_2$p[3]`). 

**SBP (<160):**
The same as before but separately for males and females... 

**Effect modification:**
When looking at the estimates and comparing it to the crude estimate ($\text{gender}_{crude}$) it appears that elevated systolic blood pressure ($\text{sbp}_{>160}$) has a universal, strong and statistically significant effect on the cardiovascular disease; no matter the gender. However, this is not telling if there is effect modification or not... 


## 2. MULTIPLE LOGISTISK REGRESSION MODEL

### 2.1 ASSOCIATION BETWEEN SYSTOLIC BLOOD PRESSURE ($\text{sbp}$), THE RISK OF GETTING CARDIOVASCULAR DISEASE ($\text{chd}$) ADJUSTING FOR GENDERS ($\text{sex}$)

To evaluate how the two genders ($\text{sex}$) effect the association a multiple logistic regression model, in which $\text{sex}$ is included, is made:
$$\text{chd} = \beta_\text{0} + \beta_\text{1} \cdot \text{sbp} + \beta_\text{2} \cdot \text{sex}$$
This model is compared to *the simple logistic model* in which $\text{sex}$ isn't included

```{r regression_mult, message=FALSE, warning=FALSE, include=FALSE}
table_3 <- list(
  "Model" = glm(chd ~ sbp + sex, data = chd, family = binomial) |> parameters(exponentiate = TRUE) |> as.data.frame()
  ) |> 
  bind_rows(.id = "Table 3: Multiple logistic regression") |> 
  select("Table 3: Multiple logistic regression" = Parameter, 
         Estimate = Coefficient, CI_low, CI_high, p) |> 
  mutate(
    across(c(Estimate, CI_low, CI_high), ~ round(.x, 2)),
    p = case_when(
      p < 0.001 ~ "< 0.001",
      p < 0.05  ~ "< 0.05",
      TRUE      ~ as.character(round(p, 3))
    )
  )
```
```{r table_mult, echo=FALSE, message=FALSE, warning=FALSE}
library(stargazer)

## Models
model_mult   <- glm(chd ~ sbp + sex, data = chd, family = binomial)

## Transforming from log-odds to odds
OR_mult <- exp(coef(model_mult))
CI_mult <- exp(confint(model_mult))
p_mult  <- summary(model_mult)$coefficients[, 4]

## Table
stargazer(model, model_mult,
          type = "text",                  
          title = "Table 3: Association between cardiovascular disease, SBP (ref. = >160 mmHg) & gender (ref. = male)",
          column.labels = c("Simple model", "Multiple model"),
          coef = list(OR_vector, OR_mult),   
          ci = TRUE,
          ci.custom = list(CI_vector, CI_mult),  
          p = list(p_values, p_mult),         
          covariate.labels = c("SBP (<160)", "Gender (female)", "Intercept"),
          digits = 2,
          star.cutoffs = c(0.05, 0.01, 0.001),
          keep.stat = c("n")
)
```
**INTERPRETATION**

When looking at the results the risk when having elevated systolic blood pressure ($\text{sbp}_{>160}$), the gender-specific OR (`r table_2$Estimate[1]` & `r table_2$Estimate[3]`) is higher than the crude model (`r table_1$Estimate[1]`)

This phenomenon is called confounding, more specific Simpson's Paradox, and is characterized by;

1. $\text{sex}$ is strongly associated to developing cardiovascular disease $\text{chd}$ e.g. $\text{sex}_{female}$ has significantly lower odds than $\text{sex}_{male}$ independently to $\text{sbp}$.  

2. $\text{sex}$ is associated to the SBP-distribution; the genders has different distribution of high SBP $\text{sbp}_{>160}$ and normal SBP $\text{sbp}_{<160}$. 
*In conclusion* when not adjusting for genders ($\text{sex}$), the effect of high SBP ($\text{sbp}_{>160}$) is underestimated because $\text{sex}_{male}$, with a higher overall risk, and $\text{sex}_{female}$, with a lower overall risk, are mixed together.


##### 2.1.1. CONFOUNDING

To better evaluated if gender ($\text{sex}$) is a confounder to the association, a plot is constructed:

```{r confounding_plot, echo=FALSE, message=FALSE, warning=FALSE}
library(patchwork)

## Collecting the prediction values
plot_data <- plot_predictions(model_mult,
                              condition = list("sbp", "sex"),
                              type = "response", 
                              df = model_mult$df.residual,
                              draw = FALSE)

## Constructing the plot
ggplot(plot_data, 
       aes(x = sbp, y = estimate, 
           color = sex, group = sex)) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), 
                width = 0.1, linewidth = 0.8, position = position_dodge(0.3)) +
  geom_point(size = 4, position = position_dodge(0.3)) +
  scale_color_manual(values = c("Female" = "##E6A15C", "Male" = "##1F3A52")) +
  labs(title = "Figure 1: Predicted Probability of Cardiovascular Disease (chd)",
       subtitle = "Based on Systolic Blood Pressure and Gender",
       x = "Systolic Blood Pressure (mmHg)",
       y = "Predicted Probability of chd",
       color = "Gender") +
  theme_minimal(base_size = 11) +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(face = "italic", size = 12)
  ) +
  plot_annotation(caption = "Dataset: Framingham (chd)")
```

*Figure 1 illustrates that males ($\text{sex}_{male}$) have a significantly higher predicted probability of developing cardiovascular disease $\text{chd}$ compared to females ($\text{sex}_{female}$) across both SBP-strata*, as indicated by the non-overlapping 95% CIs between genders within each group.

However, the relative increase in risk from normal to elevated SBP develops in a similar parallel fashion for both men and women. This parallel trend visually confirms that sex ($\text{sex}$) acts as a confounder because it shifts the baseline risk rather than modifying the effect of SBP itself.


### 2.2 ASSOCIATION BETWEEN SBP ($\text{sbp}$) & THE RISK OF GETTING CARDIOVASCULAR DISEASE ($\text{chd}$) ADJUSTING FOR GENDERS ($\text{sex}$) & AGE ($\text{age}$)

**Model**
$$\text{chd} = \beta_\text{0} + \beta_\text{1} \cdot \text{sbp} + \beta_\text{2} \cdot \text{sex} + \beta_\text{3} \cdot \text{age}$$
```{r regression_resp_sex_age, message=FALSE, warning=FALSE, include=FALSE}
table_4 <- list(
  "Model" = glm(chd ~ sbp + sex + age, data = chd, family = binomial) |> parameters(exponentiate = TRUE) |> as.data.frame()
  ) |> 
  bind_rows(.id = "Table 4: Multiple logistic regression") |> 
  select("Table 4: Multiple logistic regression" = Parameter, 
         Estimate = Coefficient, CI_low, CI_high, p) |> 
  mutate(
    across(c(Estimate, CI_low, CI_high), ~ round(.x, 2)),
    p = case_when(
      p < 0.001 ~ "< 0.001",
      p < 0.05  ~ "< 0.05",
      TRUE      ~ as.character(round(p, 3))
    )
  )
```
```{r table_resp_sex_age, echo=FALSE, message=FALSE, warning=FALSE}
library(stargazer)

## Models
model_mult1   <- glm(chd ~ sbp + sex + age, data = chd, family = binomial)

## Transforming from log-odds to odds
OR_mult1 <- exp(coef(model_mult1))
CI_mult1 <- exp(confint(model_mult1))
p_mult1  <- summary(model_mult1)$coefficients[, 4]

## Table
stargazer(model_mult1, 
          type = "text",
          title = "Table 3: Association between cardiovascular disease & SBP (ref. = >160 mmHg) adjusting for gender (ref. = male) & age",
          coef = list(OR_mult1),
          ci = TRUE,
          ci.custom = list(CI_mult1),
          p = list(p_mult1),
          covariate.labels = c("SBP (>160)", "Gender (female)", "Age (year)", "Intercept"),
          digits = 2,
          star.cutoffs = c(0.05, 0.01, 0.001),
          keep.stat = c("n")
)
```
**INTERPRETATION**

**Reference group:**
A male ($\text{sex}_{male}$) with normal blood pressure ($\text{sbp}_{<160}$) with an age ($\text{age}$) of 0 years old. 

**Intercept:**
The odds for getting cardiovascular disease ($\text{chd}$) for the reference group..

**SBP (>160):**
When adjusting for gender ($\text{sex}$) and age ($\text{age}$), individuals with high SBP ($\text{sbp}_{>160}$) is `r table_4$Odds[2]` (`r table_4$CI_low[2]`; `r table_4$CI_high[2]`, p `r table_4$p[2]`) times higher odds of getting cardiovascular disease ($\text{chd}$) than individuals with normal SBP ($\text{sbp}_{<160}$). 

**Gender (female):**
When adjusting for SBP ($\text{sbp}$) and age ($\text{age}$), $\text{sex}_{female}$ has `r table_4$Odds[3]` (`r table_4$CI_low[3]`; `r table_4$CI_high[3]`, p `r table_4$p[3]`) times higher odds of getting cardiovascular disease ($\text{chd}$) than $\text{sex}_{male}$.

**Age (year):**
For each year one individual gets older, the odds for evolving cardiovascular diseases ($\text{chp}$) with `r table_4$Odds[4]` (`r table_4$CI_low[4]`; `r table_4$CI_high[4]`, p `r table_4$p[4]`) when adjusting for gender ($\text{sex}$) and SBP ($\text{sbp}$) . 


##### 2.2.3 THE SPECIFIC EFFECT FOR TWO INDIVIDUALS & THE OR BETWEEN THE TWO 

a. 60-year-old men with a systolic blood pressure < 160 mmHg.

b. 50-year-old women with a systolic blood pressure > 160 mmHg.

To calculate the specific effect for the two individuals the function `hypothesis (0 = reference)`

```{r names_coefficients, message=FALSE, warning=FALSE, include=FALSE}
coefficients(model_mult1)
```
```{r hypothesis_test, echo=FALSE, message=FALSE, warning=FALSE}
## ## Hypothesis
hypotheses <- c(
  "Person A" = "`(Intercept)` + 0*`sbp>160` + 0*`sexFemale` + 60*`age` = 0",
  "Person B" = "`(Intercept)` + 1*`sbp>160` + 1*`sexFemale` + 50*`age` = 0",
  "Diff."    = "`(Intercept)` + 0*`sbp>160` + 0*`sexFemale` + 60*`age` =
                `(Intercept)` + 1*`sbp>160` + 1*`sexFemale` + 50*`age`"
)

## ## Tabel
table_hyp <- hypotheses(model_mult1,
                        hypothesis = hypotheses,
                        df = model_mult1$df.residual) |> 
  parameters(exponentiate = TRUE, digits = 1) |> 
  as.data.frame() |> 
  select("Hypothesis test" = Parameter, Odds = Coefficient, CI_low, CI_high, p) |> 
  mutate(
    across(c(Odds, CI_low, CI_high), ~ round(.x, 2)),
    p = case_when(p < 0.001 ~ "< 0.001",
                  p < 0.05  ~ "< 0.05",
                  TRUE      ~ as.character(round(p, 3)))
  )

## ## Print
table_hyp
```
**INTERPRETATION**

It's evident, that *Person A* has significant higher odds of getting cardiovascular diseases than *Person B* - however the difference  between these two individuals are NOT significant (`r table_hyp$Odds[3]` (`r table_hyp$CI_low[3]`; `r table_hyp$CI_high[3]`, p `r table_hyp$p[3]`) )


---------------
An alternative approach to generating individual-level predictions involves the `predictions`-function; this routine facilitates the estimation of both predicted odds, log-odds, and risk. 

```{r predictions_function, echo=FALSE, warning=FALSE}
table_5 <- list(
  "Person A (log-odds)" = predictions(model_mult1,
                                      newdata = datagrid(sex = c("Male"),
                                                         age = c(60), 
                                                         sbp = c("<160")), 
                                      type = "link") |> 
    parameters() |> as.data.frame(),
  "Person A (odds)" = predictions(model_mult1, 
                                  newdata = datagrid(sex = c("Male"),
                                                     age = c(60), 
                                                     sbp = c("<160")), 
                                  type = "link", transform = exp) |> 
    parameters() |> as.data.frame(),
  "Person A (risk)" = predictions(model_mult1, 
                                  newdata = datagrid(sex = c("Male"),
                                                     age = c(60), 
                                                     sbp = c("<160"))) |> 
    parameters() |> as.data.frame(),
  "Person B (log-odds)" = predictions(model_mult1,
                                      newdata = datagrid(sex = c("Female"),
                                                         age = c(50), 
                                                         sbp = c(">160")),
                                      type = "link") |>
    parameters() |> as.data.frame(),
  "Person B (odds)" = predictions(model_mult1,
                                  newdata = datagrid(sex = c("Female"),
                                                     age = c(50), 
                                                     sbp = c(">160")),
                                  type = "link", transform = exp)|>
    parameters() |> as.data.frame(),
  "Person B (risk)" = predictions(model_mult1,
                                  newdata = datagrid(sex = c("Female"),
                                                     age = c(50), 
                                                     sbp = c(">160")))|>
    parameters() |> as.data.frame()
  ) |> 
  bind_rows(.id = "Table 5: Odds, log(odds) & risk") |> 
  select("Table 5: Odds, log(odds) & risk", Estimate = Predicted, CI_low, CI_high, p) |> 
  mutate(
    across(c(Estimate, CI_low, CI_high), ~ round(.x, 2)),
    p = case_when(
      p < 0.001 ~ "< 0.001",
      p < 0.05  ~ "< 0.05",
      TRUE      ~ as.character(round(p, 3))
    )
  )
table_5

```
























