---
title: "TASK (dataset: postterm)"
author: "Amanda Wittendorf"
date: '`r Sys.Date()`'
output: html_document
---

## 1. DESKRIPTION ON THE DATASET

The dataset contains information on post-term birth for 12311 parturient women. More specifically: 

- $\text{ptd}$:     Status for post-term birth; if the child was after week 42 is defined as post-term while a child born before or after week 42 is defined as not born post-term (Yes / No)

- $\text{parity}$:  The number of which a mother has given birth; primiparous is defined as a first time mother while multiparous is defined as women who have given birth before (primipara / multipara)

- $\text{age}$:     Maternal age (year)


```{r dataset, message=FALSE, warning=FALSE, include=FALSE}
library(statistik)

post <- read.csv("postterm.csv") |> 
  mutate(ptd = factor(ptd),
         ptd = fct_relevel(ptd, "No"),
         ptd_num = as.numeric(ptd),
         parity = factor(parity),
         parity = fct_relevel(parity, "primiparous")
         )
```


## 2. SIMPLE LOGISTIC REGRESSION MODEL

### 2.1 ASSOCIATION BETWEEN STATUS FOR POST-TERM BIRTH ($\text{ptd}$) & MATERNAL AGE $\text{age}$.

**The model**
$$\text{ptd} = \beta_\text{0} + \beta_\text{1} \cdot \text{age}$$
```{r regression_ptd_age, message=FALSE, warning=FALSE, include=FALSE}
table_1 <- list(
  "Model" = glm(ptd ~ age, data = post, family = binomial) |> parameters(exponentiate = TRUE) |> as.data.frame()
  ) |> 
  bind_rows(.id = "Table 1: Simple logistic regression") |> 
  select("Table 1: Simple logistic regression" = Parameter, 
         Odds = Coefficient, CI_low, CI_high, p) |> 
  mutate(
    across(c(Odds, CI_low, CI_high), ~ round(.x, 2)),
    p = case_when(
      p < 0.001 ~ "< 0.001",
      p < 0.05  ~ "< 0.05",
      TRUE      ~ as.character(round(p, 3))
    )
  )
```
```{r table_ptd_age, echo=FALSE, message=FALSE, warning=FALSE}
library(stargazer)

## Models
model   <- glm(ptd ~ age, data = post, family = binomial)

## Transforming from log-odds to odds
OR_vector <- exp(coef(model))
CI_vector <- exp(confint(model))
p_values  <- summary(model)$coefficients[, 4]

## Table
stargazer(model, 
          type = "text",                  
          coef = list(OR_vector), 
          title = "Table 1: Association between post-term birth (ref. = No) & maternal age",
          ci = TRUE,                      
          ci.custom = list(CI_vector),    
          p = list(p_values),           
          dep.var.labels = "Post-term birth (ptd)",
          covariate.labels = c("Maternal age (year)", "Intercept"),
          digits = 2,
          star.cutoffs = c(0.05, 0.01, 0.001), 
          keep.stat = c("n")
)
```
**INTREPETATION**

**Intercept:** 
The statistically significant odds for having post-term birth ($\text{ptd}$) for mothers who has a maternal age at 0 years ($\text{age}=0$) - this is not of biological meaning because of the maternal age. 

**Maternal age (years)**
For each year the maternal age increases, the odds for having a post-term birth will increases significant with `r table_1$Odds[2]` (`r table_1$CI_low[2]`; `r table_1$CI_high[2]`, p`r table_1$p[2]`). 


#### 2.1.1 Calculate OR for a specific age difference of 10 years. 

To calculate this, the `hypotheses()`-function is used without including the intercept. This is because we are evaluating the relative effect of a 10-year age difference rather than an absolute value. Conversely, if we were to predict the specific odds or probability for an individual case, the intercept would be required.

```{r names_coefficients, message=FALSE, warning=FALSE, include=FALSE}
coefficients(model)
```
```{r hypothesis_test, echo=FALSE, message=FALSE, warning=FALSE}
## ## Hypothesis
hypotheses <- c(
  "Age difference, 10 years" = "10*age = 0"
)

## ## Tabel
table_hyp <- hypotheses(model,
                        hypothesis = hypotheses,
                        df = model$df.residual) |> 
  parameters(exponentiate = TRUE, digits = 1) |> 
  as.data.frame() |> 
  select("Hypothesis test" = Parameter, OR = Coefficient, CI_low, CI_high, p) |> 
  mutate(
    across(c(OR, CI_low, CI_high), ~ round(.x, 2)),
    p = case_when(p < 0.001 ~ "< 0.001",
                  p < 0.05  ~ "< 0.05",
                  TRUE      ~ as.character(round(p, 3)))
  )

## ## Print
table_hyp
```
**INTERPRETATION**

The odds of having a post-term birth is significantly higher when comparing mothers with an age difference of 10 years; `r table_hyp$OR[2]` (`r table_hyp$CI_low[2]`; `r table_hyp$CI_high[2]`, p`r table_hyp$p[2]`). 

The p-value remains the same when comparing the exposure ($\text{age}$) and the predicted value ($\text{age}_{diff(10)}$) because both tests evaluate the same underlying hypothesis: *whether maternal age is associated with post-term birth* 


## 3. MULTIPLE LOGISTIC REGRESSION MODEL

### 3.1. ASSOCIATION BETWEEN STATUS FOR POST-TERM BIRTH ($\text{ptd}$) & MATERNAL AGE $\text{age}$ ADJUSTED FOR PARITY ($\text{parity}$)

**The model**
$$\text{ptd} = \beta_\text{0} + \beta_\text{1} \cdot \text{age} + \beta_\text{2} \cdot \text{parity}$$
```{r regression_multi, message=FALSE, warning=FALSE, include=FALSE}
table_2 <- list(
  "Model" = glm(ptd ~ age + parity, data = post, family = binomial) |> parameters(exponentiate = TRUE) |> as.data.frame()
  ) |> 
  bind_rows(.id = "Table 2: Multiple logistic regression") |> 
  select("Table 2: Multiple logistic regression" = Parameter, 
         Odds = Coefficient, CI_low, CI_high, p) |> 
  mutate(
    across(c(Odds, CI_low, CI_high), ~ round(.x, 2)),
    p = case_when(
      p < 0.001 ~ "< 0.001",
      p < 0.05  ~ "< 0.05",
      TRUE      ~ as.character(round(p, 3))
    )
  )
```
```{r table_multi, echo=FALSE, message=FALSE, warning=FALSE}
library(stargazer)

## Models
model_mult   <- glm(ptd ~ age + parity, data = post, family = binomial)

## Transforming from log-odds to odds
OR_mult <- exp(coef(model_mult))
CI_mult <- exp(confint(model_mult))
p_mult  <- summary(model_mult)$coefficients[, 4]

## Table
stargazer(model, model_mult, 
          type = "text",                  
          coef = list(OR_vector, OR_mult), 
          title = "Table 2: Comparing the simple & multiple model for the association between post-term birth (ref. = No) & maternal age, adjusted for parity",
          column.labels = c("Simple model", "Multiple model"),
          ci = TRUE,                      
          ci.custom = list(CI_vector, CI_mult),    
          p = list(p_values, p_mult),           
          dep.var.labels = "Post-term birth (ptd)",
          covariate.labels = c("Maternal age (year)", "Parity status (ref. = primiparous)", "Intercept"),
          digits = 2,
          star.cutoffs = c(0.05, 0.01, 0.001), 
          keep.stat = c("n")
)
```

#### 3.1.1. CONFOUNDING 

The purpose of this section is to evaluate if *($\text{parity}$) is a confounder for the association*. 

```{r confounding_plot, echo=FALSE, message=FALSE, warning=FALSE}
## Collecting the prediction values (Husk at bruge din opdaterede model her)
plot_data <- plot_predictions(model_mult,
                              condition = list("age", "parity"),
                              type = "response",
                              df = model_mult$df.residual,
                              draw = FALSE)

## Constructing the plot
ggplot(plot_data, 
       aes(x = age, y = estimate, 
           color = parity, fill = parity, group = parity)) + 
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.15, color = NA) + 
  geom_line(linewidth = 1) + 
  
  scale_color_manual(
    values = c("primipara" = "#E6A15C", "multipara" = "#1F3A52"),
    labels = c("primipara" = "Primiparous (=1)", "multipara" = "Multiparous (>1)")) + 
  
  scale_fill_manual(
    values = c("primipara" = "#E6A15C", "multipara" = "#1F3A52"),
    labels = c("primipara" = "Primiparous (=1)", "multipara" = "Multiparous (>1)")) + 
  
  labs(title = "Figure 1: Predicted Probability of post-term birth (prd)",
       subtitle = "Based on Maternal Age & Parity",
       x = "Maternal age (year)",
       y = "Predicted Probability of prd",
       color = "Parity", fill = "Parity", 
       caption = "Dataset: postterm.csv") + 
  
  theme_minimal(base_size = 11) +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(face = "italic", size = 12)
  )

```

**INTERPRETATION**

The slope estimate doesn't changed when comparing the crude model with the adjusted one; see $\text{Table 2,}$ $\text{OR}_{age}$. This indicates that $\text{parity}$ isn't a confounder for the association between post-term birth ($\text{pdr}$) and maternal age ($\text{age}$). This effect is called *independent main effects; the parity groups and the maternal age have their own independent and statistically significant risk factors for the association to post-term births*.


#### 3.1.2. EFFECT MODIFICATION

To evaluate if parity status have an effect modification on the association, an interactive term is included in the model:

**The model**
$$\text{ptd} = \beta_\text{0} + \beta_\text{1} \cdot \text{age} + \beta_\text{2} \cdot \text{parity} + \beta_\text{3} \cdot \text{parity} \cdot \text{age}$$

```{r regression_multi1, message=FALSE, warning=FALSE, include=FALSE}
table_3 <- list(
  "Model" = glm(ptd ~ age * parity, data = post, family = binomial) |> parameters(exponentiate = TRUE) |> as.data.frame()
  ) |> 
  bind_rows(.id = "Table 3: Multiple logistic regression with interaction") |> 
  select("Table 3: Multiple logistic regression with interaction" = Parameter, 
         Odds = Coefficient, CI_low, CI_high, p) |> 
  mutate(
    across(c(Odds, CI_low, CI_high), ~ round(.x, 2)),
    p = case_when(
      p < 0.001 ~ "< 0.001",
      p < 0.05  ~ "< 0.05",
      TRUE      ~ as.character(round(p, 3))
    )
  )
table_3
```
```{r table_multi1, echo=FALSE, message=FALSE, warning=FALSE}
library(stargazer)

## Models
model_mult1   <- glm(ptd ~ age * parity, data = post, family = binomial)

## Transforming from log-odds to odds
OR_mult1 <- exp(coef(model_mult1))
CI_mult1 <- exp(confint(model_mult1))
p_mult1  <- summary(model_mult1)$coefficients[, 4]

## Table
stargazer(model, model_mult1, 
          type = "text",                  
          coef = list(OR_vector, OR_mult1), 
          title = "Table 3: Comparing the simple & mult1iple model for the association between post-term birth (ref. = No) & maternal age, with a interactive term between age and parity",
          column.labels = c("Simple model", "Mult1iple model"),
          ci = TRUE,                      
          ci.custom = list(CI_vector, CI_mult1),    
          p = list(p_values, p_mult1),           
          dep.var.labels = "Post-term birth (ptd)",
          covariate.labels = c("Maternal age (year)", "Parity status (ref. = primiparous)", "Interaction (age:parity)", "Intercept"),
          digits = 2,
          star.cutoffs = c(0.05, 0.01, 0.001), 
          keep.stat = c("n")
)
```
```{r effect_modification_plot, echo=FALSE, message=FALSE, warning=FALSE}
## Collecting the prediction values (Husk at bruge din opdaterede model her)
plot_data <- plot_predictions(model_mult1,
                              condition = list("age", "parity"),
                              type = "response",
                              df = model_mult1$df.residual,
                              draw = FALSE)

## Constructing the plot
ggplot(plot_data, 
       aes(x = age, y = estimate, 
           color = parity, fill = parity, group = parity)) + 
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.15, color = NA) + 
  geom_line(linewidth = 1) + 
  
  scale_color_manual(
    values = c("primipara" = "#E6A15C", "multipara" = "#1F3A52"),
    labels = c("primipara" = "Primiparous (=1)", "multipara" = "Multiparous (>1)")) + 
  
  scale_fill_manual(
    values = c("primipara" = "#E6A15C", "multipara" = "#1F3A52"),
    labels = c("primipara" = "Primiparous (=1)", "multipara" = "Multiparous (>1)")) + 
  
  labs(title = "Figure 1: Predicted Probability of post-term birth (prd)",
       subtitle = "Based on Maternal Age & Parity",
       x = "Maternal age (year)",
       y = "Predicted Probability of prd",
       color = "Parity", fill = "Parity", 
       caption = "Dataset: postterm.csv") + 
  
  theme_minimal(base_size = 11) +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(face = "italic", size = 12)
  )

```

**INTERPRETATION**

Table 3 shows that there is no effect modification between parity status and the association between maternal age and post-term birth. The interaction term is not statistically significant, with an $\text{OR}_{interactive}$, `r table_3$Odds[4]` (`r table_3$CI_low[4]`; `r table_3$CI_high[4]`, p `r table_3$p[4]`). 


#### 3.1.3. The effect of maternal age ($\text{age}$) for each of the two parity groups ($\text{parity}$).

```{r names_coefficients_mult, message=FALSE, warning=FALSE, include=FALSE}
coefficients(model_mult1)
```
```{r hypothesis_test_mult, echo=FALSE, message=FALSE, warning=FALSE}
## ## Hypothesis
hypotheses <- c(
  "Primiparaous (=1)" = "age = 0",
  "Multiparaous (>1)" = "age + `age:parityprimipara` = 0" 
)

## ## Tabel
table_hyp <- hypotheses(model_mult1,
                        hypothesis = hypotheses,
                        df = model_mult1$df.residual) |> 
  parameters(exponentiate = TRUE, digits = 1) |> 
  as.data.frame() |> 
  select("Hypothesis test" = Parameter, OR = Coefficient, CI_low, CI_high, p) |> 
  mutate(
    across(c(OR, CI_low, CI_high), ~ round(.x, 2)),
    p = case_when(p < 0.001 ~ "< 0.001",
                  p < 0.05  ~ "< 0.05",
                  TRUE      ~ as.character(round(p, 3)))
  )

## ## Print
table_hyp
```
**INTERPRETATION**

**Primiparous (=1):**
For first time mothers the odds increases significantly for post-term births with `r table_hyp$OR[1]` (`r table_hyp$CI_low[1]`; `r table_hyp$CI_high[1]`, p `r table_hyp$p[1]`) for each year the mother gets older. 

**Multiparous (>1):**
For multiparous mothers the odds increases significantly for post-term births with  `r table_hyp$OR[2]` (`r table_hyp$CI_low[2]`; `r table_hyp$CI_high[2]`, p `r table_hyp$p[2]`) for each year the mother gets older. 

The small difference between the two parity groups is a reflection of the non-significant effect modification shown in *table 3*. 



#### 3.1.4. The specific effect for two women

Unlike in *Section 2.1.1*, the intercept is included here because we are calculating the specific odds for two individual profiles.

a. A 35-year-old multiparous woman.

b. A 30-year-old primiparous woman.

```{r hypothesis_test1, echo=FALSE, message=FALSE, warning=FALSE}
## ## Hypothesis
hypotheses <- c(
  "Person A" = "`(Intercept)` + 35*`age` + 1*`parityprimipara` + 35*`age:parityprimipara` = 0",
  "Person B" = "`(Intercept)` + 30*`age` + 0*`parityprimipara` + 30*`age:parityprimipara` = 0",
  "Diff." = "`(Intercept)` + 35*`age` + 1*`parityprimipara` + 35*`age:parityprimipara` =
            `(Intercept)` + 30*`age` + 0*`parityprimipara` + 30*`age:parityprimipara`"
)

## ## Tabel
table_hyp <- hypotheses(model_mult1,
                        hypothesis = hypotheses,
                        df = model_mult1$df.residual) |> 
  parameters(exponentiate = TRUE, digits = 1) |> 
  as.data.frame() |> 
  select("Hypothesis test" = Parameter, OR = Coefficient, CI_low, CI_high, p) |> 
  mutate(
    across(c(OR, CI_low, CI_high), ~ round(.x, 2)),
    p = case_when(p < 0.001 ~ "< 0.001",
                  p < 0.05  ~ "< 0.05",
                  TRUE      ~ as.character(round(p, 3)))
  )

## ## Print
table_hyp
```
**INTERPRETATION**

*Person A has a significant higher odds* for post-term births ($\text{pdr}$) than Person B; this difference is, however, not statistically significant (p `r table_hyp$p[3]`).





