---
title: "TASK (dataset: perulung)"
author: "Amanda Wittendorf"
date: "`r Sys.Date()`"
output: html_document
---

# 1. DESKRIPTION ON THE DATASET

The dataset contains measurements from 636 Peruvian children. More specifically: 

- $\text{FEV1}$: Forceret Ekspiratorisk Volumen in the 1. secound 

- $\text{height}$: childrens height (cm) 

- $\text{age}$: childrens age (year) 

- $\text{sex}$: gender (0 = girls / 1 = boys)

- $\text{respsymptoms}$: measured respiratory symptoms within the past year (0 = no / 1 = yes)

```{r dataset, message=FALSE, include=FALSE}
library(statistik)

perulung <- read.csv("perulung.csv") |> 
  mutate(sex = factor(sex,
                      levels = c(0, 1), 
                      labels = c("Girls", "Boys")),
         sex = fct_relevel(sex, "Boys"),
         respsymptoms = factor(respsymptoms,
                               levels = c(0, 1),
                               labels = c("No", "Yes")),
         respsymptoms = fct_relevel(respsymptoms, "Yes"))
```


# 2. SIMPLE LOGISTIC REGRESSION MODEL

## 2.1. ASSOCIATION BETWEEN RESPIRATORY SYMPTOMS ($\text{respsymptoms}$) & GENDERS ($\text{sex}$)

```{r regression_resp_sex, message=FALSE, warning=FALSE, include=FALSE}
table_1 <- list(
  "Model" = glm(respsymptoms ~ sex, data = perulung, family = binomial) |> parameters(exponentiate = TRUE) |> as.data.frame()
  ) |> 
  bind_rows(.id = "Table 1: Simple logistic regression") |> 
  select("Table 1: Simple logistic regression" = Parameter, 
         Estimate = Coefficient, CI_low, CI_high, p) |> 
  mutate(
    across(c(Estimate, CI_low, CI_high), ~ round(.x, 2)),
    p = case_when(
      p < 0.001 ~ "< 0.001",
      p < 0.05  ~ "< 0.05",
      TRUE      ~ as.character(round(p, 3))
    )
  )
```
```{r table_resp_sex, echo=FALSE, message=FALSE, warning=FALSE}
library(stargazer)

# Models
model   <- glm(respsymptoms ~ sex, data = perulung, family = binomial)

# Transforming from log-odds to odds
OR_vector <- exp(coef(model))
CI_vector <- exp(confint(model))
p_values  <- summary(model)$coefficients[, 4]

# Table
stargazer(model, 
          type = "text",                  
          coef = list(OR_vector), 
          title = "Table 1: Association between respiratory symptoms (ref. = no) and sex (ref. = boys)",
          ci = TRUE,                      
          ci.custom = list(CI_vector),    
          p = list(p_values),             
          d$ep.var.labels = "Respiratory symptoms",
          covariate.labels = c("Gender (Female)", "Intercept"),
          digits = 2,
          star.cutoffs = c(0.05, 0.01, 0.001), 
          keep.stat = c("n")
)
```
**INTREPETATION**

The estimates is interpreted as odds/odds ratio (OR). 

**Intercept:** 
The statistically significant odds for $\text{gender}_{boys}$ (reference group) with respiratory symptoms is `r table_1$Mean[1]` (`r table_1$CI_low[1]`; `r table_1$CI_high[1]`, p `r table_1$p[1]`) which means, that every time one boy appear without symptoms, `r table_1$Mean[1]` boys will appear with symptoms.  

**Gender (female):**
The odds ratio (OR) for $\text{gender}_{girls}$ comparing to $\text{gender}_{boys}$ with both having respiratory symptoms is `r table_1$Mean[2]` (`r table_1$CI_low[2]`; `r table_1$CI_high[2]`, p `r table_1$p[2]`). However, this is not significant. 

--------------
To calculate the specific effect for $\text{gender}_{girls}$ the function `hypothesis` is used: 

```{r names_coefficients, echo=FALSE}
coefficients(model)
```
```{r hypothesis_test, echo=FALSE, message=FALSE, warning=FALSE}
# Hypothesis
hypotesis <- c(
  "Odds for girls" = "`(Intercept)` + sexGirls = 0",
  "Odds for boys" = "`(Intercept)` = 0",
  "OR (girls vs. boys)"  = "sexGirls = 0" 
)

# Tabel
table_hyp <- hypotheses(model, 
                      hypothesis = hypotesis, 
                      df = model$df.residual) |>
  parameters(exponentiate = TRUE, digits = 1) |> as.data.frame() |>
  
  select("The effect of respiratory symptoms " = Parameter, Mean = Coefficient, CI_low, CI_high, p) |> 
  mutate(
    across(c(Mean, CI_low, CI_high), ~ round(.x, 2)),
         p = case_when(p < 0.001 ~ "< 0.001",
                       p < 0.05  ~ "< 0.05",
                       TRUE      ~ as.character(round(p, 3)))
         )

# Print
table_hyp
```

**INTERPRETATION**

The last effect, the effect of respiratory symptoms for $\text{gender}_{girls}$, is now calculated to be statistically significant `r table_hyp$Mean[1]` (`r table_hyp$CI_low[1]`; `r table_hyp$CI_high[1]`, p `r table_hyp$p[1]`). 
























